Semi-supervised low-rank logistic regression for high-dimensional neuroimaging data

### 2015-05-30 paper_nips/paper.tex
done first pass

\paragraph{Optimization.}
The common objective was \suggestremove{approximated} \suggestadd{optimized}
% XXX why 'approximated' ?
in the neuroimaging data
by \suggestremove{updating parameters by derivates}
\suggestadd{gradient descent in the parameters}
% XXX derivates ???
of the 
semi-supervised low-rank logistic regression.
The required gradients are easily obtained by using the chain rule to
backpropagate error derivatives through the linear network.
As solver, we chose \textit{rmsprop} \cite{rmsprop},
a mini-batch version of rprop.
This procedure dictates an adaptive learning rate
for each model parameter by
scaled gradients from a running average.
Gradient normalization by rmsprop
is known to effectively exploit curvature information.
We opted for a small batch size of $100$, given the high degree of
redundancy in $X_{rest}$ and $X_{task}$.
The matrix parameters were initalized by Gaussian random values multiplied
by a gain of $0.004$. The bias parameters were initalized to zero.
With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
The normalization factor and the update rule for $\theta$
are given by

\begin{eqnarray}
  \begin{split}
    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
\\
\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
  \end{split}
\end{eqnarray}

where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
$0.9$ to deemphasize the magnitude of the gradient.
Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
%
Note that we have also experimented with other solvers
(stochastic gradient descent, adadelta, and adagrad) but found that
rmsprop converged faster and with higher generalization performance.

\paragraph{Hints.}
In fact, the constraint by a rest-data autoencoder qualifies as a
\textit{hint}
rather than regularization in a strict sense \cite{abu1994hints}.
Its purpose is not to prevent overfitting but to introduce
prior knowledge on known properties of the unknown target function $f$.
Rather than only relying on input-output pairs in the learning process,
we thus narrow our hypothesis set to the biologically most plausible solutions.
That is, we reduce the search space in a way that
is compatible with the expected representation of BOLD activity signals.

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the high-dimensional neuroimaging data 
\cite{abrah14}
and
\textit{theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
\url{http://github.com/anonymous/anonymous}.
% url{http://github.com/banilo/nips2015}.

@@

\begin{figure}
% XXX: figure not readable
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/semisup.png}
\end{centering}
\caption{\textbf{Relation between support recovery and out-of-dataset performance}
\textit{Upper row:} the fit-for-purpose of the classification models
is quantified by linear correlation between the
z-scored average task image and the z-scored weights from
ordinary (\textit{red}) and present (\textit{blue colors}) logistic
regression ($\lambda=0.25, 0.50, 0.75, 1.00$).
\textit{Lower row:} ordinary logistic regression is applied to the
ARCHI dataset after projection into the latent dimensions learned from
the HCP task data.
\textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
}
}
\end{figure}

>>>>

\paragraph{Optimization.}
The common objective was \suggestremove{approximated} \suggestadd{optimized}
% XXX why 'approximated' ?
in the neuroimaging data
by \suggestremove{updating parameters by derivates}
\suggestadd{gradient descent in the parameters}
% XXX derivates ???
of the 
semi-supervised low-rank logistic regression.
The required gradients are easily obtained by using the chain rule to
backpropagate error derivatives through the linear network.
As solver, we chose \textit{rmsprop} \cite{rmsprop},
a mini-batch version of rprop.
This procedure dictates an adaptive learning rate
for each model parameter by
scaled gradients from a running average.
Gradient normalization by rmsprop
is known to effectively exploit curvature information.
We opted for a small batch size of $100$, given the high degree of
redundancy in $X_{rest}$ and $X_{task}$.
The matrix parameters were initalized by Gaussian random values multiplied
by a gain of $0.004$. The bias parameters were initalized to zero.
With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
The normalization factor and the update rule for $\theta$
are given by
% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
% XXX done in an alternating manner. The whole structure relies on this, by
% XXX construction it is vulnerable to optimizing too much towards labels or
% XXX rest at a time. I also identify this as one of the major weak points
% XXX of the method, so it needs to be stated in a way that is not easily
% XXX attackable.
% XXX Maybe we could also talk less about RMSprop.

\begin{eqnarray}
  \begin{split}
    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
\\
\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
  \end{split}
\end{eqnarray}
% XXX Michael: Why, in the formula, do we have two different notations for
% XXX partial derivatives? \partial and \nabla

% XXX Michael: Probably remove the explanations following and just state values
where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
$0.9$ to deemphasize the magnitude of the gradient.
Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
%
Note that we have also experimented with other solvers
(stochastic gradient descent, adadelta, and adagrad) but found that
rmsprop converged faster and with higher generalization performance.

\paragraph{Hints.}
In fact, the constraint by a rest-data autoencoder qualifies as a
\textit{hint}
rather than regularization in a strict sense \cite{abu1994hints}.
Its purpose is not to prevent overfitting but to introduce
% XXX Michael: I use f earlier up, is this f here necessary?
prior knowledge on known properties of the unknown target function $f$.
Rather than only relying on input-output pairs in the learning process,
we thus narrow our hypothesis set to the biologically most plausible solutions.
That is, we reduce the search space in a way that
is compatible with the expected representation of BOLD activity signals.

\paragraph{Implementation.}
The analyses were performed in Python.
We used \textit{nilearn} to handle
the high-dimensional neuroimaging data 
\cite{abrah14}
and
\textit{Theano} for automatic, numerically stable
differentiation of symbolic computation graphs
\cite{bastien2012theano, bergstra2010theano}.
All Python scripts that generated the results are
accessible online for reproducibility and reuse
\url{http://github.com/anonymous/anonymous}.
% url{http://github.com/banilo/nips2015}.

@@

\begin{figure}
% XXX: figure not readable
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/semisup.png}
\end{centering}
\caption{\textbf{Relation between support recovery and out-of-dataset performance}
\textit{Upper row:} the fit-for-purpose of the classification models
is quantified by linear correlation between the
z-scored average task image and the z-scored weights from
ordinary (\textit{red}) and present (\textit{blue colors}) logistic
regression ($\lambda=0.25, 0.50, 0.75, 1.00$).
\textit{Lower row:} ordinary logistic regression is applied to the
ARCHI dataset after projection into the latent dimensions learned from
the HCP task data.
\textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
{\color{blue} Michael: Does this also happen when you double the task data?
i.e. rest:task 3:1, but duplicate the task data, yielding essentially 3:2?
I assume this will work, which again points to the instability of the network architecture}}
}
\end{figure}
